---
output: md_document
---

```{r message=FALSE, warning=FALSE}
library(MESS)
library(splines)
library(dplyr)
library(survival)
library(quantreg)

source("R/fit_sens.R")
source("R/fit_spec.R")
source("R/predict_sens.R")
source("R/predict_spec.R")
source("R/lsurvROC.R")
source("R/process_dat.R")
source("R/plot_coef_lsurvROC.R")
source("R/plot_coef_lsurvROC.R")
source("R/get_AUC.R")
source("R/monotone_ROC.R")
source("R/get_ROC.R")
source("R/plot_ROC.R")
source("R/plot_curve_lsurvROC.R")
```

## Introduction

We propose a flexible regression framework to evaluate the prognostic capability of longitudinal biomarker with survival outcome accounting for the covariate and biomaker measurement time. Here is an example using a simulated dataset to estimate the biomarker threshold at a specified specificity level along with the corresponding sensitivity, and to construct a covariate- and measurement time-specific ROC curve.


## Read in long and short data
```{r message=FALSE, warning=FALSE}
dat.long <- read.csv("Data/reg_data_sim_long.csv")
dat.short <- read.csv("Data/reg_data_sim_short.csv")

```

## Estimate time-varying coefficients

 - dat.long: long format data, each subject may have multiple biomarker records
 - dat.short: short format data, each subject only has one record, including observed event time, event indicator, covariates (eg. Z, Zcont in this example)
 - cutoff.type.basis: type of basis function to estimate biomarker threshold, (FP: Fractional polynomial, BS: B Splines, linear: linear basis)
 - sens.type.basis: type of basis function to estimate sensitivity level, (FP: Fractional polynomial, BS: B Splines, linear: linear basis)
 - covariate1: a vector of covariate names to estimate the biomarker threshold 
 - covariate2: a vector of covariate names to estimate the sensitivity level
 - tau: target specificity level, could be a single value or a set of values between 0 and 1 
 - time.window: a time window for biomarker evaluation
 - nResap: number of perturbation resampling
 
 
```{r}
model.FP1 = lsurvROC(dat.long = dat.long, 
                            dat.short = dat.short, 
                            cutoff.type.basis = "FP", 
                            sens.type.basis = "FP", 
                            covariate1 = c("Z", "Zcont"), 
                            covariate2 = c("Z", "Zcont"), 
                            tau = 0.8, 
                            time.window = 1,
                            nResap = 100,
                            show_plots = FALSE)


par(mfrow=c(1,3))
model.FP1$plots$cutoff()
model.FP1$plots$sensitivity()


```

## Conditional ROC curve 

```{r}
#estimate thresholds and sensitivity levels
model.FP2 <- lsurvROC(dat.long = example_data$data.long, 
                          dat.short = example_data$data.short,
                          cutoff.type.basis = "FP",
                          sens.type.basis = "FP", 
                          covariate1 = c("Z", "Zcont"), 
                          covariate2 = c("Z", "Zcont"), 
                          tau = seq(0.1, 0.9, 0.05), 
                          time.window = 1, 
                          nResap = 50, 
                          show_plots = FALSE
)
#get monotoned ROC curve and AUC value
ROC.results <- plot_lsurvROC(model = model.FP2$models, 
                              my.newdat = data.frame(vtime = 0.5, Z = 1, Zcont = 0.25), 
                              tau = seq(0.1, 0.9, 0.05),
                              basis = "FP", 
                              tol = 1e3, 
                              add = FALSE,
                              col = "black", 
                              lty = 1)

ROC.results$AUC
ROC.results$AUC.sd
```


